---
title: "Comparing methods to adjust cases by number of PCR tests"
output: html_notebook
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(modelr)

source(here("functions", "retrieve_tests_cases_region.R"))
```

Import data and filter Metropolitan Region:
```{r}
examenes_casos_rm <- 
  retrieve_tests_cases_region() %>% 
  filter(Region == "Metropolitana")
```

## Model 1

One linear regression per week (n=7), with different slope and intercept.
```{r}
df_adj_cases_model1 <- 
  examenes_casos_rm %>% 
  nest(data = c(-Semana)) %>% 
  mutate(reg_casos = map(data, ~lm(N_Casos ~ N_Tests, data = .)),
         adjusted_cases = map_dbl(reg_casos, 
                                  ~predict(., newdata = tibble(N_Tests = 10000)))) 

df_adj_cases_model1 %>% 
  select(Semana, adjusted_cases)
```

## Model 2

One linear regression for all the data (n=255), with fixed effects per week (different intercepts, but the same slope).
```{r}
model2 <- 
  lm(N_Casos ~ factor(Semana) + N_Tests, data = examenes_casos_rm)

df_adj_cases_model2 <- 
  examenes_casos_rm %>% 
  mutate(N_Tests = 10000) %>% 
  add_predictions(model2, var = "adjusted_cases") %>%
  group_by(Semana) %>% 
  summarise(adjusted_cases = mean(adjusted_cases))

df_adj_cases_model2
```

## Model 3 

Same as model 2, but using a logarithmic function to consider decreasing marginal confirmed cases as testing capacity is increased:
```{r}
model3 <- 
  lm(N_Casos ~ factor(Semana) + log(N_Tests), data = examenes_casos_rm)

df_adj_cases_model3 <- 
  examenes_casos_rm %>% 
  mutate(N_Tests = 10000) %>% 
  add_predictions(model3, var = "adjusted_cases") %>%
  group_by(Semana) %>% 
  summarise(adjusted_cases = mean(adjusted_cases))

df_adj_cases_model3
```

# Comparing methods

Looking at the curves of adjusted cases:
```{r}
df_adj_cases_all_models <- 
  df_adj_cases_model1 %>% 
  rename(model1 = adjusted_cases) %>% 
  left_join(df_adj_cases_model2 %>% 
              rename(model2 = adjusted_cases), by = "Semana") %>% 
  left_join(df_adj_cases_model3 %>% 
              rename(model3 = adjusted_cases), by = "Semana") %>% 
  pivot_longer(
    cols = starts_with("model"),
    names_to = "model",
    values_to = "adjusted_cases"
  )

ggplot(df_adj_cases_all_models,
       aes(Semana, adjusted_cases, color = model)) +
  geom_line()
```

Plotting the predicted cases curves in a Cases ~ Tests scatter plot.
```{r}
examenes_casos_rm_filt <- 
  examenes_casos_rm %>% 
  filter(Semana %in% c(c(31, 35, 40, 49, 51)))

examenes_casos_rm_filt %>% 
  ggplot(aes(N_Tests, N_Casos, color = factor(Semana))) +
  geom_point()
```

Creating grid of tests, weeks, and cases:
```{r}
df1 <- 
examenes_casos_rm %>% 
  modelr::data_grid(Semana, N_Tests) %>% 
  filter(Semana %in% c(c(31, 35, 40, 49, 51))) %>% 
  left_join(df_adj_cases_model1 %>% 
              select(Semana, reg_casos), by = "Semana") %>% 
  add_predictions(model2, var = "model2") %>% 
  add_predictions(model3, var = "model3") %>% 
  mutate(model1 = map2_dbl(reg_casos, N_Tests, 
                           ~predict(object = .x, newdata = tibble(N_Tests = .y)))) 
  
  
df1 %>% 
  pivot_longer(
    cols = starts_with("model"),
    names_to = "model",
    values_to = "adjusted_cases"
  ) %>% 
  ggplot(aes(N_Tests, color = factor(Semana))) +
  geom_line(aes(y = adjusted_cases)) +
  facet_wrap(~model) +
  labs(color = "Week") +
  geom_point(data = examenes_casos_rm_filt,
             aes(N_Tests, N_Casos))
```

Intersecting slopes are problematic because the questions I want to answer is "are we better off in week X compared to week Y?"... and it can't be true for a week to have less virus prevalence than other for certain levels of testing, but more for others (testing levels don't change virus prevalence).

## Caso de semana 51
```{r}
df_sem51 <- 
  examenes_casos_rm %>% 
  filter(Semana == 51)

bind_rows(
  `Including yesterday` = df_sem51,
  `Excluding yesterday` = df_sem51 %>% 
    filter(Fecha != lubridate::ymd(20201218)),
  .id = "type_df"
) %>% 
  ggplot(aes(N_Tests, N_Casos)) +
  geom_point(aes(color = type_df)) +
  geom_smooth(aes(color = type_df), method = "lm", se = FALSE, fullrange=TRUE) +
  expand_limits(x = 5000) +
  labs(color = "Data",
       y = "Confirmed cases",
       x = "Tests") +
  annotate("label",
           label = "Yesterday\n(dec. 18)",
           x = 17000,
           y = 620,
           color = "#00BFC4")
```

